---
title: "RALLF Forage Quality"
subtitle: "preliminary report"
date: "2023 February"
author: "Jesse Puka-Beals"
output:
  # word_document:
    # reference_docx: "word_style.docx"
  # pdf_document:
    # toc: true
editor_options: 
  chunk_output_type: console
# format: 
#   pdf: 
#     colorlinks: false
#     include-in-header: 
#       text: |
bibliography: rallf.bib
csl: asa-cssa-sssa.csl
---

### To do

Validate the predicted forage quality parameters with wet chemistry. When we predicted NDFD for IWG in FIG the values were reasonable (50%), but for alfalfa they are unreasonable (16%). For example, @Grev2017 observed NDFD range of 33-45% for a 30 day cutting interval.

Investigate the 3rd cut in st paul 2022.  

\newpage

### Expected results

Alfalfa varieties (HarvXtra) will have lower lignin content and higher NDFD than conventional varieties. The RFV should be higher for HarvXtra varieties.

Alfalfa with lower intervals between cuttings (35-day vs 45-day ) will have higher forage quality (RFV, CP).

### Discussion topics

*RFV vs. RFQ*

If we are only making comparisons among alfalfa cuttings, let's use RFV not RFQ. RFQ relies on NDFD, which is much harder to predict with NIRS. There is lot of variability in NDFD in wet chemsity due to variability in rumen fluid composition from donor cows, this variability extends into the NIRS database for NDFD, which reduces the accuracy of the NIRS predictions. RFQ is useful when comparing between mixtures of legumes and grasses, but since we are only making comparisons between alfalfa cuttings, RFV allows us to rely on just ADF and NDF, which NIRS performs better at predicting.

*NIRS predictions*

Do we trust the predicted values of the NIRS for this experiment?

Is the hay equation sufficient for making comparisons between HarvXtra vs. conventional alfalfa forage quality?

Why is the RFV prediction so high for alfalfa? 

Why is neutral detergent digestibility so low? We need to be skeptical of RFQ so long as we are skeptical of NDFD.

Why is variance so low in the NIRS prediction? 

### Data availability

\href{https://github.com/jpukabeals/rallf}{Click Me for Github Repository} [*https://github.com/jpukabeals/rallf*](https://github.com/jpukabeals/rallf){.uri}

```{r, echo=F, include=F}
knitr::opts_chunk$set(
  echo = F,
  warning = F, 
  message = F,
  comment = NA
)

library(tidyverse)
library(tidyr)

rm(list = ls())

theme_set(theme_bw())

```

```{r, include=F}

source("rallf_nir_addInTreatmentInfo.R")
source("functions_nir.R")
# rm(key1,key2,key3)

```

\newpage

# Data we have

```{r}
key3 %>% 
  left_join(key2) %>% 
  left_join(key1) %>% 
  left_join(dat) %>% 
  filter(harvest_code <= 14) %>%  # we did not harvest beyond more than 4 cuts in 2022
  group_by(
    year,site,intensity,cut, harvest_code
  ) %>%
  summarise(n_missing=sum(is.na(CP))) %>% 
  mutate(n = 24-n_missing,.before=n_missing) %>% 
  arrange(harvest_code) %>% 
  dplyr::select(-harvest_code) %>%
  knitr::kable(
    caption = "NIRS data"
  )

key3 %>% 
  left_join(key2) %>% 
  left_join(key1) %>% 
  left_join(dat) %>% 
  filter(harvest_code <= 14) %>%  # we did not harvest beyond more than 4 cuts in 2022
  group_by(
    year,site,intensity,cut, harvest_code
  ) %>%
  summarise(n_missing=sum(is.na(CP))) %>% 
  mutate(n = 24-n_missing,.before=n_missing) %>% 
  arrange(harvest_code) %>% 
  dplyr::select(-harvest_code) %>%
  filter(n_missing != 0) %>% 
  knitr::kable(
    caption = "NIRS data"
  )

```

As seen in tables, we have missing $24+24+1 = 85$ data points. The two sets of 24 data points are for the first cut at rosemount in 2022, which have not been located as of 15Mar2023.

\newpage

```{r}

dat %>% 
  # colnames()
  dplyr::select(code,year,plot,
                site,intensity,cut,variety,rep,
                CP,ADF,NDF,NDFD,rfv,
                rfq.legume) -> dat1

```

# Predicting forage quality parameters with NIRS

We are predicting alfalfa forage quality parameters using near infrared scanning with a generalist equation designed for all hay. Some labs use an alfalfa equation and some labs have different alfalfa equations for conventional vs. reduced lignin alfalfa. All predictions should be validated with wet chemistry, though NIRS predictions without recent validation can still be useful in estimating relative differences between treatments.

Soil contamination can also cause issues with forage quality predictions.

Expected forage quality parameters for alfalfa were easy to obtain for CP ADF NDF and RFV, they were not easy to obtain for RFQ. In general, I found that RFQ can range from 50 to 250, with most data between 100 and 200 and an average of 150.

```{r Dairyland predicted values}
c(
  "CP",
  "ADF",
  "NDF",
  "Lignin"
) -> prmtr

c(
  20,
  31,
  38,
  6
) -> mdn

c(
  14,
  20,
  35,
  4
) -> mn

c(
  26,
  43,
  52,
  8
) -> mx

tibble(
  prmtr,
  mdn,
  mn,
  mx
) %>% 
  mutate(
    source="dairyland labs",
    .before=prmtr
  ) %>% 
  mutate(range = paste(mn,mx,sep = "-")) %>% 
  rename(
    median = mdn,
    parameter = prmtr
  ) %>% 
  dplyr::select(-c(mn,mx,source)) %>% 
  rename_all(str_to_title) %>% 
  knitr::kable(
    caption = "Expected forage quality ranges from Dairyland Labs"
  )

```

```{r predicted ranges aggregated}

# https://docs.google.com/spreadsheets/d/1AqVcEwv3PlHVcR5Vv5UqUliUcqOZj6dNMvK-FAArONg/edit#gid=1646351822

read.csv(
  "alfalfa_forageQuality_expectedValues.csv"
) -> dat_qual


dat_qual %>% 
  rename(
    "Grade" = quality.category,
  ) %>% 
  rename_all(str_to_title) %>% 
  rename_with(
    str_to_upper,
    4:8
  ) %>% 
  knitr::kable(
    caption= "Expected forage quality ranges from UMN, SDSU, Univeristy of Georgia and Dairyland"
  )

```

```{r observed vs expected}

dat_qual %>% 
  dplyr::filter(stage=="all") %>% 
  dplyr::select(-c(maturity,stage,quality.category)) %>% 
  rename_all(toupper) %>% 
  mutate(NDFD = "40-60") %>% 
  mutate(type = "expected") -> d1


dat1 %>% 
  summarise(
    across(
      c(CP,ADF,NDF,NDFD,rfv,rfq.legume),
      mean
    )
  ) %>% 
  rename(
    RFQ = rfq.legume,
    RFV = rfv
  ) %>% 
  mutate_all(
    ~round(.x,0)
  ) %>% 
  mutate(
    type = "observed"
  ) %>% 
  mutate_all(as.character) %>% 
  bind_rows(d1) %>% 
  relocate(type,.before=CP) -> d2 

# d2 %>% 
#   knitr::kable()

c(
  20,
  30,
  40,
  50,
  150,
  150
) -> expctd

as_tibble(cbind(nms = names(d2), t(d2))) %>% 
  rename(
    Observed = V2,
    `Expected Range` = V3,
    `Forage Quality Parameter` = nms
  ) %>% 
  slice(-1) %>% 
  bind_cols(expctd) %>% 
  rename(Expected = "...4") %>% 
  relocate(Expected,.after=Observed) %>% 
  knitr::kable(
    caption = "Observed vs. expected forage quality in RALLF"
  )

```

```{r predicted ranges from Grev paper}
read.csv(
  "grev2017.csv"
) -> grev

grev %>% 
  pivot_longer(
    cols = 4:9
    # cols = starts_with("CP"),
  ) %>% 
  mutate(fq_component = str_split(name,"_",simplify = T)[,1],
         site = str_split(name,"_",simplify = T)[,2]) %>% 
  dplyr::select(-name) %>% 
  mutate(value = value/10) %>% 
  group_by(fq_component,type) %>% 
  summarise(mean=mean(value),
            max = max(value),
            min = min(value)) %>%
  mutate(
    across(
      2:4,
      ~format(round(.x),nsmall=0)
    )
  ) %>%
  mutate(range = paste0(min,"-",max)) %>% 
  mutate(meanrange = paste0(mean," (",range,")")) %>% 
  dplyr::select(-c(max,min,mean,range)) %>% 
  pivot_wider(
    names_from = fq_component,
    values_from = c(meanrange)
  ) %>% 
  rename(Type = type) %>% 
  mutate_all(str_to_title) %>% 
  knitr::kable(
    caption = "Forage quality parameters reported by Grev 2017 on alfalfa stands in St Paul and Rosemount planted in spring and cut on 30 day harvest intervals over two years. Mean and (range) are presented."
  )

```


```{r FQ observed vs grev}

grev %>% 
  pivot_longer(
    cols = 4:9
    # cols = starts_with("CP"),
  ) %>% 
  mutate(fq_component = str_split(name,"_",simplify = T)[,1],
         site = str_split(name,"_",simplify = T)[,2]) %>% 
  dplyr::select(-name) %>% 
  mutate(value = value/10) %>% 
  group_by(fq_component,type) %>% 
  summarise(mean=round(mean(value)))  %>%
  pivot_wider(
    names_from = fq_component,
    values_from = c(mean)
  ) %>% 
  rename(Type = type) %>% 
  mutate_all(str_to_title) %>% 
  mutate(source = "Grev et al. 2017")->dfgrev

dat1 %>% 
  mutate(variety = fct_recode(variety,
                              `reduced lignin` = "HX3",
                              `reduced lignin` = "HX4",
                              `reduced lignin` = "HX5",
                              conventional = "RR3",
                              conventional = "RR4",
                              conventional = "RR5")) %>%
  mutate(variety = str_to_title(variety)) %>% 
  group_by(variety) %>% 
  rename_with(
    .cols = c(1:8),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  rename(
    Type = Variety
  ) %>% 
  dplyr::select(Type,CP,NDF,NDFD) %>% 
  mutate(source = "Observed") %>% 
  mutate_all(as.character) %>% 
  bind_rows(dfgrev) %>% 
  relocate(source,.before=Type) %>% 
  knitr::kable(
    caption = "Forage quality comparison between Grev et al 2017 and RALLF"
  )

```


\newpage

# Summary tables

```{r FQ by rep}

dat1 %>% 
  group_by(rep) %>% 
  rename_with(
    .cols = c(1:8),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by rep"
  )

```

```{r FQ by site}
dat1 %>% 
  group_by(site) %>% 
  rename_with(
    .cols = c(1:8),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by site"
  )

```

```{r FQ by year}
dat1 %>% 
  group_by(year) %>% 
  rename_with(
    .cols = c(1:8),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  # colnames()
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by year"
  )
```

```{r FQ by cut}
dat1 %>% 
  group_by(cut) %>% 
  rename_with(
    .cols = c(1:8),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by cutting"
  )
```

```{r FQ by cutting intensity}
dat1 %>% 
  group_by(intensity) %>% 
  rename_with(
    .cols = c(1:8),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by harvest schedule"
  )

```

\newpage

```{r FQ by variety}

dat1 %>% 
  group_by(variety) %>% 
  rename_with(
    .cols = c(1:8),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by variety"
  )

```


\newpage

# Histograms

```{r RFQ}

dat1 %>% 
  # colnames()
  ggplot(aes(rfq.legume)) +
  stat_bin() +
  labs(
    x="RFQ"
  )

```

We expect an RFQ of around 150, we observed an RFQ of `r round(mean(dat1$rfq.legume),0)`

```{r}
dat1 %>% 
  # colnames()
  ggplot(aes(rfv)) +
  stat_bin() +
  labs(
    x="RFV"
  )
```

We expect an RFV of around 150, we observed an RFV of `r round(mean(dat1$rfv),0)`. Some samples were predicted to have an RFV greater than 300!

```{r}
dat1 %>% 
  # colnames()
  ggplot(aes(rfv,
             fill = factor(cut))) +
  stat_bin() +
  labs(
    x="RFV"
  ) +
  geom_vline(xintercept = 200) +
  facet_wrap(~site*year)
```

With the exception of Rosemount in 2022, all site-years had multiple cuts with predicted RFV over 200. First cuts tended to have the highest quality, which is expected.

```{r}
dat1 %>% 
  ggplot(aes(CP)) +
  stat_bin() 
```

We expect a crude protein of around 20%, we observed a CP of `r round(mean(dat1$CP),0)`

```{r}
dat1 %>% 
  ggplot(aes(ADF)) +
  stat_bin()

```

We expect an ADF of around 30%, we observed an ADF of `r round(mean(dat1$ADF),0)`

```{r}
dat1 %>% 
  ggplot(aes(NDF)) +
  stat_bin()

```

We expect an NDF of around 40%, we observed an NDF of `r round(mean(dat1$NDF),0)`

```{r}
dat1 %>% 
  ggplot(aes(NDFD)) +
  stat_bin()

```

For neutral detergent digestibility over 48 hours, we expect values near 50%, we predicted values of `r round(mean(dat1$NDFD),0)`

\newpage

# Analysis

## Relative feed value

Since relative feed value and relative forage quality are strongly correlated, the results of the analysis tend to be the same.

```{r echo=T}

dat1 %>% 
  lm(
    rfv~site*year*variety*intensity*cut,.
  ) %>% 
  anova() %>%
  filter(`Pr(>F)` < 0.05)
# year:intensity:cut
# site:year:cut
# site:year:intensity

# no effect from variety

```

\newpage

```{r}
dat1 %>% 
  ggplot(aes(cut,rfv,
             col=intensity,
             fill=intensity)) +
  stat_summary(
    geom="bar",
    position=position_dodge(.5),
    width = .5,
    col=1
  ) +
  stat_summary(
    geom="errorbar",
    position=position_dodge(.5),
    width = .3,
    col=1
  ) +
  facet_wrap(~site*year) +
  labs(y="RFV")
```

Similar pattern across years and sites; RFV is similar for the first cut and then the second cut the RFV is lower for the 45-day cuts than the 35-day cuts. The third cut at St Paul in 2022 is unexpected, what could explain this?

Why is variance so low? Standard error bars are barely visible. 

```{r checking variance, echo=T, include=F}
# is their psuedoreplication/duplication happening causing variance to be low?
# this could be a mistake I made during a left_join()
dat1 %>% 
  filter(
    site=="rosemount" &
      year == "2021" &
      intensity == "45-day" &
      cut=="2"
  ) %>% 
  distinct(rfq.legume) %>% 
  .$rfq.legume 
# note that there are no duplicated values of RFQ

```

Let's look at standard deviation instead of standard error

```{r}
dat1 %>%
  ggplot(aes(cut, rfv,
             col = intensity,
             fill = intensity)) +
  stat_summary(
    geom = "bar",
    position = position_dodge(.5),
    width = .5,
    col = 1
  ) +
  stat_summary(
    fun.data = mean_sdl,
    geom = "errorbar",
    position = position_dodge(.5),
    width = .3,
    col = 1
  ) +
  facet_wrap( ~ site * year) +
  labs(y = "RFV + Standard deviation")
```

There is simply a large difference between standard error and standard deviation, likely because n() is larger than usual (n=24) so standard error really reduces the visual representation of variance.

\newpage

## Crude protein

```{r echo=T}

dat1 %>% 
  lm(
    CP~site*year*variety*intensity*cut,.
  ) %>% 
  anova() %>%
  filter(`Pr(>F)` < 0.05)
# 4-way interaction

# no effect from variety

```

\newpage

```{r}
dat1 %>% 
  ggplot(aes(cut,CP,
             col=intensity,
             fill=intensity)) +
  stat_summary(
    geom="bar",
    position = position_dodge(.5),
    width = .5,
    col=1) +
  facet_wrap(~site*year) +
  labs(y="CP")
```

Similar pattern across years and sites, same as RFQ; CP is similar for the first cut and then the second cut the CP is lower for the 45-day cuts than the 35-day cuts.

Same as with RFQ, summarizing with standard error shows much less variation around means than standard deviation.

\newpage

# Diving deeper

## Relative feed value and ADF

RFV is the only parameter where we are outside of our expected range. We expect around 150 and we are averaging 200. We have some RFV values over 300.

This may be due to an issue with the equation

Let's look at the RFV equation I used

```{r}
print(calc.rfq.rfv.legume)
```

So I used ...

Digestible dry matter\
$DDM.rfv = 88.9-(0.779*ADF)$

Dry matter intake\
$DMI.rfv = 120/NDF$

Relative Forage Value\
$RFV = (88.9-(0.779*ADF))*(120/NDF)/1.29$

These equations are correct.

```{r testing RFV equation, include=F}

# low range of fiber
ADF <- 27
NDF <- 34
# RFV was 185

ADF <- 29
NDF <- 36
# RFV was 171

# expected values
ADF <- 30
NDF <- 40
# RFV=150

# observed values
ADF <- 24
NDF <- 33
# RFV = 200

# ratio of ADF:NDF weird?
33/24
# 1:1.3
40/30
# 1:1.3

# the ratio of ADF to NDF is what we expect
# for 1ADF there's 1.3 NDF

(88.9-(0.779*ADF))*(120/NDF)/1.29
```

Let's look at the samples with the highest RFV values

```{r}
dat1 %>% 
  arrange(desc(rfv)) %>% 
  slice(1:5) %>% 
  dplyr::select(
    site,year,intensity,cut,variety,rep,
    CP,ADF,NDF,NDFD,rfv,rfq.legume
  ) %>% 
  # print()
  mutate(
    across(
      7:12,
      ~round(.x,0)
    )
  ) %>% 
  knitr::kable(
    caption = "The samples with the highest estiamted relative feed value"
  )
  
```

The forage samples with the highest RFV are from st paul in 2022. The ADF is 16% and the NDF is 20%. A forage sample with these values would have a relative feed value of 355, so the equation is not incorrect. The question is whether alfalfa can be this high quality.

Can an alfalfa sample be 16% ADF and 20% NDF?

It's notable that the crude protein is also very high, 29%, where we would normally expect 20%. This supports the idea that this was young vegetative leafy alfalfa.

It seems very possible that our predicted ADF and NDF values simply need a bias adjustment. If this is the case, they can still be used to determine relative differences between treatments, but will need a wet chemistry calibration before they can be used to determine the actual ADF or NDF values.

\newpage

## NDFD

Neutral detergent fiber digestibility estimates the amount of digestion that will occur within the rumen and this varies by ruminant. The most common durations are 24 hours for cattle, 30 hours for sheep goats and horses, 48 hours for swine. There are also very long digestions (72, 120 or 240), which are meant to determine how much fiber is indigestible.

We predict neutral detergent fiber digestibility over 48 hours because this is used for the RFQ equation [@Undersander2002]. We expect an NDFD range of 30-50% for alfalfa based on data from Dairyland and Forage Genetics International, but NDFD can range up to 70%[@Hoffman2001].

```{r expected lignin and NDFD values}

c(
  "HarvXtra",
  "Conventional"
) ->alf_var

c(
  "35-day",
  "45-day"
) -> intensity


c(
  5,4.5,6.5,5.7
) -> lgnn

c(
  40,49,35,43
) -> ndfd_p

expand_grid(
  alf_var,
  rev(intensity)
) %>% 
  bind_cols(
    lgnn,ndfd_p
  ) %>% 
  rename(
    Variety = alf_var,
    Intensity = `rev(intensity)`,
    `Lignin (%)` = ...3,
    `NDFD (%)` = ...4
  ) %>% 
  knitr::kable(
    caption = "Expected lignin and NDFD values of different varieties and harvest schedules"
  )
# https://www.foragegenetics.com/FGI/media/Images/PDF/FGI_Comparison_Between_HarvXtra_Alfalfa_and_Varieties_Selected_for_High-Quality_White_Paper_1.pdf
```

```{r, include=F}
dat1 %>% 
  # group_by(variety,intensity) %>% 
  # summarise(NDFD = mean(NDFD)) %>% 
  mutate(Variety = str_sub(variety,1,2)) %>% 
  group_by(Variety,intensity) %>% 
  summarise(NDFD_48hr = mean(NDFD),
            sd = sd(NDFD)) %>% 
  # glimpse()
  mutate(
    across(
      c(NDFD_48hr,sd),
      .fns = ~round(.x,1)
    )
  ) %>% 
  mutate(
    Variety = factor(Variety),
    Variety = fct_recode(Variety,
                         HarvXtra = "HX",
                         Conventional = "RR")
  ) %>%  
  knitr::kable(
    caption = "Predicted NDFD mean and standard deviation observed in RALLF between HarvXtra (HX) and conventional (RR) alfalfa"
  )
```

```{r}

dat1 %>% 
  # group_by(variety,intensity) %>% 
  # summarise(NDFD = mean(NDFD)) %>% 
  mutate(Variety = str_sub(variety,1,2)) %>% 
  group_by(Variety,intensity) %>% 
  summarise(NDFD_48hr = mean(NDFD),
            sd = sd(NDFD)) %>% 
  # glimpse()
  mutate(
    across(
      c(NDFD_48hr,sd),
      .fns = ~round(.x,1)
    )
  ) %>% 
  mutate(
    Variety = factor(Variety),
    Variety = fct_recode(Variety,
                         HarvXtra = "HX",
                         Conventional = "RR")
  ) %>% 
  dplyr::select(-sd) %>% 
  mutate(type = "NDFD_observed") %>% 
  rename(`NDFD (%)` = NDFD_48hr,
         Intensity = intensity) -> dm1


expand_grid(
  alf_var,
  rev(intensity)
) %>% 
  bind_cols(
    lgnn,ndfd_p
  ) %>% 
  rename(
    Variety = alf_var,
    Intensity = `rev(intensity)`,
    `Lignin (%)` = ...3,
    `NDFD (%)` = ...4
  ) %>% 
  mutate(type = "NDFD_expected") %>% 
  dplyr::select(-`Lignin (%)`) %>% 
  bind_rows(dm1) %>% 
  relocate(type,.before=Variety) %>% 
  # print()
  pivot_wider(
    names_from=type,
    values_from = `NDFD (%)`
  ) %>% 
  knitr::kable(
    caption = "Predicted NDFD values did not match with expected values"
  )

```

We observed around 15% NDFD, which very low.

We expected as time between cuttings decreased (35-day harvest schedule), NDFD would increase, however we observed a decrease in NDFD for both the HarvXtra and Conventional varieties

We expected greater NDFD in HarvXtra vs conventional, but observed lower NDFD.

One possible explanation is that our NDFD prediction needs a new equation for alfalfa NDFD as it differs from grass NDFD.

Another explanation is that we need a different equation for HarvXtra alfalfa vs. conventional alfalfa.

The predicted NIRS values for NDFD should be viewed with skepticism, and it likely cannot be fixed with a bias. 

```{r}

dat1 %>% 
  # glimpse()
  summarise(
    max(NDFD),
    min(NDFD),
    median(NDFD)
  ) %>% 
  mutate_all(
    ~round(.x,0)
  ) %>% 
  knitr::kable(
    caption = "Neutral detergent digestibility ranged from 4 to 31%. This is very low and should be validated with wet chemistry"
  )
```

Does the low NDFD values make the RFQ values lower, causing the RFQ and RFV to not be well correlated?

```{r}
library(corrplot)

# dat1 %>% 
#   dplyr::select(rfv,rfq.legume) %>% 
#   cor() %>% 
#   corrplot()
# ok, RFQ and RFV are very positively correlated

dat1 %>% 
  dplyr::select(rfv,rfq.legume, ADF, NDF,CP,NDFD) %>% 
  cor() %>% 
  corrplot()
# As NDF and ADF increase, so does NDFD, and RFV and RFQ goes down. 
  
```

Yes, RFV and RFQ are strongly positively correlated as they should be.

Other forage quality parameters are correlated as expected with the exception of NDFD.

As forages mature, CP should decrease and NDFD should decrease, therefore they should be positively correlated, but instead they are negatively correlated. Furthermore, as NDFD increases RFQ should increase, but they are also negatively correlated.

\newpage

## RFQ

RFQ should be positively correlated with RFV and have a close relationship (slope near 1)

$$
RFQ = 1.2*RFV - 32.4
$$
[@Undersander2002]

```{r}

lm(
  rfq.legume~rfv,dat
)

dat %>% 
  ggplot(aes(rfv,rfq.legume)) +
  geom_point() +
  geom_smooth(method = "lm",se=F) +
  labs(
    caption = "RFQ = 0.7*RFV -1.4"
  )
```

We observe a relationship of $RFV = 0.7*RFV - 1.4$

This indicates there is an issue with how we are estimating some of the forage quality parameters that comprise RFV and RFQ

```{r include=F}
RFQ.pred <- function(RFV){
  1.2*RFV - 32.4
}
RFQ.pred(187)
```

#### ADF NDF underestimation

We observed low ADF and NDF values, but they were within an acceptable range. The RFV prediction of 205, however, was outside the expected range of 100-200. When we compared our NIRS prediction with Europe, we estimated lower ADF and NDF from the same sorghum sudangrass. If we simply increased ADF by `r 2/32*100`% and NDF by `r 5/60*100`%, the RFV goes from 205 to 187 and is within the acceptable range. 

```{r include=F}
dat %>% 
mutate(
  ADF=ADF*1.065,
  NDF=NDF*1.083,
  # NDFD=NDFD*2.8,
           NFC=100-((0.93*NDF)+CP+EE+Ash),
           NDFn=NDF*0.93, 
           NDFDp=22.7+0.664*NDFD,
           TDN.legume=(NFC*.98)+(CP*.93)+(FA*.97*2.25)+(NDFn*NDFD/100)-7,
           DMI.legume=(120/NDF) + (NDFD - 45) * .374/1350*100,
           TDN.grass=(NFC*.98)+(CP*.87)+(FA*.97*2.25)+(NDFn*NDFDp/100)-10,
           DMI.grass=(-2.318)+(.442*CP)-(.01*CP^2)-(.0638*TDN.grass)+(.000922*TDN.grass^2)+
             (.18*ADF)-(0.00196*ADF^2)-(0.00529*CP*ADF),
           rfq.legume=DMI.legume*TDN.legume/1.23,
           DDM.rfv = 88.9-(.779*ADF),
           DMI.rfv = 120/NDF,
           rfv= DDM.rfv*DMI.rfv/1.29) %>% 
  summarise(rfq = mean(rfq.legume), rfv = mean(rfv))
```

While increasing the ADF and NDF values brings down the RFV to an acceptable range, it also decreases the RFQ from 139 to 121. The relationship of $RFQ = 0.7*RFV-1.4$ remains...

#### NDFD underestimation

```{r include = F}
range(dat$NDFD)
mean(dat$NDFD)
50/16
```

Our NDFD range is also low. We expect a range of 30 to 70 [@Hoffman2001, @Undersander2016], but we observe a range of 4 to 31 with a mean of 16. We should have a mean of 50. 

Let's see how the RFV-RFQ relationship changes when we shift the NDFD mean to 50 $NDFD*3.125$ and keep our adjustment of ADF $ADF*1.0625$ and NDF $NDF*1.083$. 


```{r, include=F}
print(calc.rfq.rfv)

dat %>% 
mutate(
  ADF=ADF*1.065,
  NDF=NDF*1.083,
  NDFD=NDFD*3.125,
           NFC=100-((0.93*NDF)+CP+EE+Ash),
           NDFn=NDF*0.93, 
           NDFDp=22.7+0.664*NDFD,
           TDN.legume=(NFC*.98)+(CP*.93)+(FA*.97*2.25)+(NDFn*NDFD/100)-7,
           DMI.legume=(120/NDF) + (NDFD - 45) * .374/1350*100,
           TDN.grass=(NFC*.98)+(CP*.87)+(FA*.97*2.25)+(NDFn*NDFDp/100)-10,
           DMI.grass=(-2.318)+(.442*CP)-(.01*CP^2)-(.0638*TDN.grass)+(.000922*TDN.grass^2)+
             (.18*ADF)-(0.00196*ADF^2)-(0.00529*CP*ADF),
           rfq.legume=DMI.legume*TDN.legume/1.23,
           DDM.rfv = 88.9-(.779*ADF),
           DMI.rfv = 120/NDF,
           rfv= DDM.rfv*DMI.rfv/1.29) %>% 
  # summarise(rfq = mean(rfq.legume), rfv = mean(rfv))
  # lm(rfq.legume~rfv,.)
  ggplot(aes(rfv,rfq.legume)) +
  geom_point() +
  geom_smooth(method = "lm", se=F) +
  labs(
    caption = "After adjusting ADF NDF and NDFD"
  )

```

With an RFV of 187, we would expect an RFQ of 192 and we predict 200. Much better! By multiplying NDFD by 3.125, we are still not fixing the issue of NDFD not being positively correlated with CP, indicating a larger problem with it's prediction. More importantly, the linear relationship is ever worse now at $RFQ = 0.34*RFV + 193$

At what NDFD value could we achieve the slope closest to 1.2*RFV while keeping the ADF and NDF adjustment?

```{r include=F}
dat %>% 
mutate(
  ADF=ADF*1.065,
  NDF=NDF*1.083,
  NDFD=NDFD*2,
  EE=3,
  CP = CP*1.25,
           NFC=100-((0.93*NDF)+CP+EE+Ash),
           NDFn=NDF*0.93, 
           NDFDp=22.7+0.664*NDFD,
           TDN.legume=(NFC*.98)+(CP*.93)+(FA*.97*2.25)+(NDFn*NDFD/100)-7,
           DMI.legume=(120/NDF) + (NDFD - 45) * .374/1350*100,
           TDN.grass=(NFC*.98)+(CP*.87)+(FA*.97*2.25)+(NDFn*NDFDp/100)-10,
           DMI.grass=(-2.318)+(.442*CP)-(.01*CP^2)-(.0638*TDN.grass)+(.000922*TDN.grass^2)+
             (.18*ADF)-(0.00196*ADF^2)-(0.00529*CP*ADF),
           rfq.legume=DMI.legume*TDN.legume/1.23,
           DDM.rfv = 88.9-(.779*ADF),
           DMI.rfv = 120/NDF,
           rfv= DDM.rfv*DMI.rfv/1.29) %>% 
  # summarise(rfq = mean(rfq.legume), rfv = mean(rfv))
  lm(rfq.legume~rfv,.)
  # ggplot(aes(rfv,rfq.legume)) +
  # geom_point() +
  # geom_smooth(method = "lm", se=F) +
  # labs(
  #   caption = "After adjusting ADF NDF and NDFD"
  # )
```

We cannot achieve a slope of $RFQ = 1.2RFV$ by simply adjusting NDFD. 

What would it take? It's not clear but NDFD needs to be addressed. Simply multiplying NDFD is too crude and we likely need to improve our equation for NDFD prediction. Adjustments of ADF and NDF shift both RFQ and RFV, adjustments of CP and EE do not seem to change relationship greatly.  

Another step that could be taken would be to note the mahalanoibis distances for NDFD prediction for RALLF. We would expect them to be above 5, indicating a warning. 

#### Extractable ether

Extractable ether in our equation is set at 2.05 as a constant. But according to @Undersander2016, extractable ether ranges from 1 to 4 and averages at 3 for alfalfa/grass forage.

Adjusting extractable ether does not have a large effect on predicted RFQ or RFV, so it doesn't really matter. Just a heads up. 

\newpage

# References
